---
title: "Final Project EDA"
format: pdf
editor: visual
author: "Lindsay Kowal & Kathleen Bacigalupi"
---

## Final Project Check In

We are using the 2022 Environmental Justice Index dataset published by the [CDC](https://www.atsdr.cdc.gov/placeandhealth/eji/index.html) with data from the U.S. Census Bureau, the U.S. Environmental Protection Agency, the U.S. Mine Safety and Health Administration, and the U.S. Centers for Disease Control and Prevention. The data is filtered to Massachusetts (1474 observations) where each row is a census tract.

Question: Which pollutants correspond with higher rates of asthma in Massachusetts? Does this differ by poverty rates? (/percentage of people living under 200% of the federal poverty level?) 

Our response variable will be the percentage of individuals that have asthma (EP_ASTHMA). Our explanatory variables will be the percentage below 200% poverty (EP_POV200), the annual mean days above O3 standard (E_OZONE), the annual mean days above PM 2.5 regulatory standard (E_PM), and the ambient concentrations of diesel (E_DSLPM). 

**Data Dictionary for EDA**

<https://eji.cdc.gov/Documents/Data/2022/EJI_2022_Data_Dictionary_508.pdf>

**Data to Download**

<https://www.atsdr.cdc.gov/placeandhealth/eji/eji-data-download.html>

```{r}
us <- read.csv("United States.csv") |>
  dplyr::filter(StateDesc == "Massachusetts") 

mass |>
  dplyr::filter(!is.na(E_OZONE & E_PM & E_DSLPM & EP_ASTHMA & EP_POV200)) |>
  dplyr::select(E_OZONE, E_PM, E_DSLPM, EP_POV200, EP_ASTHMA)

```

```{r, warning = FALSE, message = HIDE}
us |>
  dplyr::select(E_OZONE, E_PM, E_DSLPM, EP_POV200) |>
  GGally::ggpairs()
```

```{r}
library(ggplot2)
ggplot(data = us, aes(x = E_OZONE, y = EP_ASTHMA)) + 
  geom_point(position = "jitter") + 
  labs(title = "EDA Visual #2", x = "Mean Days Above Ozone Standard", y = "Percentage with Asthma") + 
  theme_minimal()
```

### 
