---
title: ""
format: pdf
editor: visual
---

• Is each included table (including table captions) fully unpacked and interpreted for the reader? (0–2 points)
• Is each included figure (including figure captions) fully unpacked and
interpreted for the reader? (0–2 points)

## **Abstract**

-   Is the abstract of the appropriate length (i.e., five to six sentences)? (0–1
    points)

-   Does it provide a succinct description of the study background/
    motivation, primary research question, and regression methods used?
    (0–2 points)

-    Does it provide a succinct and specific explanation of what the analyses
    revealed and how these results inform the broader research question?
    (0–2 points)

## Introduction

-    Is the topic of the final report clearly explained and some justification/
    motivation for the analysis provided? Is the relevance of the analysis
    clearly articulated? (0–3 points)

-    Is the organization of the background and significance logical and structured to flow from the general to the specific? (0–2 points)

-    Are the research question and goals of the analysis clearly explained? Is
    the final paragraph of the introduction a brief description of the hypothesis and goals of the paper? (0–3 points)

-   Does the introduction include (at least) 2–3 citations to previous research
    in the area to support your ideas, show what has already been done, and
    illustrate what gap you are filling? (0–2 points)

Which pollutants correspond with higher rates of asthma in Massachusetts? Does this differ by poverty rates?

## Methods

### Data and Analysis Variables

• Is the broader population of interest (i.e., the target of statistical inference) clearly identified? (0–1 point)
• Is the source of the data clearly explained, with a corresponding citation?
How was this sample collected? (0–1 points)
• Are the response and explanatory variables of interest clearly identified
and described, with levels (for categorical variables) and appropriate
units (for continuous variables) provided? (0–1 points)
• Are any important data wrangling steps (e.g., removing observations
with missing values, constructing new variables) clearly described? (0–1
points)

### Statistical Methodology

Are the statistical methods and modeling approach appropriate and logical? Are they motivated by the research question and exploratory data
analysis? (0–5 points)
• Are these methods described succinctly and in sufficient detail that someone knowledgeable could replicate the analysis and evaluate the analytic
decisions made? (0–2 points)
• Is the mathematical representation of the final population regression
model(s) provided without errors? (0–2 points)
• Is the notation used for the outcome and explanatory variable(s) in this
model clearly and succinctly defined? (0–1 points)
• Is the approach for addressing missing data (if present) briefly described?
(0–1 points)

## Results

EDA

-   Are the number of observations in the final analysis dataset (and the
    number of observations removed from analysis) clearly identified? (0–1
    points)

-   Are the distributions of the response and relevant explanatory variables,
    as well as the relationships between them, clearly and accurately described? Using appropriate summary statistics and at least one data
    visualization? (0–4 points)

### Regression Analysis & Other Statistical Results

• Are the estimates of the model parameters and their accompanying standard errors provided? (0–2 points)
• Are the relevant estimated slopes, population means, hypothesis tests,
and/or predictions (as appropriate for the stated research question and
goals) correctly interpreted in context? (0–8 points)
• Are confidence/prediction intervals provided for each of these interpreted
quantities? (0–2 points)
• Are the final results of the analysis logical and explained in a clear and
straightforward manner? Are these results clearly tied back to the research question and goals? (0–3 points)
• Are the regression assumptions and model goodness-of-fit investigated
and appropriately assessed? These items may be included in the Data
Analysis Appendix rather than the body of the final report. (0–4 points)

0.007462 no response rate, perform list wise deletion before selecting random census tracks from Massachusetts

```{r, echo = FALSE}
library(dplyr)
set.seed(5)

us <- read.csv("United States.csv") |>
  dplyr::filter(StateDesc == "Massachusetts") |>
  dplyr::filter(!is.na(E_OZONE & E_PM & E_DSLPM & EP_ASTHMA & EP_POV200)) |>
  sample_n(300)
```

```{r}
us |>
  dplyr::select(E_OZONE, E_PM, E_DSLPM, EP_POV200) |>
  GGally::ggpairs()
```

```{r}
library(ggplot2)
ggplot(aes(x = EP_ASTHMA), data = us) + 
  geom_density()

ggplot(aes(x = log(EP_ASTHMA)), data = us) + 
  geom_density()

```

```{r}
asthma1 <- lm(log(EP_ASTHMA) ~ E_OZONE + E_PM + E_DSLPM + EP_POV200, data = us)
summary(asthma1)
```

```{r}
asthma.interact <- lm(log(EP_ASTHMA) ~ E_OZONE*EP_POV200 + E_PM*EP_POV200 + E_DSLPM*EP_POV200, data = us)
summary(asthma.interact)
```

```{r}
anova(asthma1, asthma.interact)
```
